---
title: "Exercise 1: Basic Demand Estimation Exercise" 
#subtitle:
author: "Reid Falconer"
#date: "5/28/2018"
output:
  html_document:
    code_folding: show
    #css: report.css
    highlight: textmate
    number_sections: no
    theme: united
    toc: no
    toc_depth: 2
    toc_float: 
      collapsed: false
urlcolor: blue
    
---

```{r, include=FALSE}
source('scripts/ex1hints.R')
```

#### 1. Construct a price variable by dividing $ sales by unit sales (employ the `sales_$` and `sales_u` variables). Explain how to interpret this price variable (i.e. what sort of average price is this?).

```{r}
# price <- mydata$Sales_USD/mydata$Sales_U
# The constructed `price` is the average price of one unit of the item across USD.
```

#### 2. Compute the mean prices across weeks of Hellman's in Jewel and the Central Region. Are they comparable? Repeat the exercise for Kraft in Jewel and the Central Region.

```{r}
mean(na.omit(hellman_at_jewel$price))
mean(hellman_at_central$price)
```

1.11 vs 1.08 - yes they are comparable.

#### Repeat for Kraft

#### 3. Price variation: Compute the standard deviation of prices across weeks of Hellman's in Jewel and the Central Region. Is there more price variation at Jewel or in the Central Region? Why? What does this tell you upfront about your ability to estimate price elasticities with either account level data or data in a large geographic market? Repeat the exercise for Kraft in Jewel and the Central Region.

```{r}
sd(hellman_at_central$price)
sd(na.omit(hellman_at_jewel$price))

sd(kraft_at_central$price)
sd(na.omit(kraft_at_jewel$price))
```

The standard deviation for hellman at Jewel is noticeably greater than at Central. (0.067 v 0.037)

Greater standard deviation by greater aggregation. The more variation you have the more difficult to estimate underlying parameters.

In Europe you don't have a "one-price policy".

More aggregation usually leads to lower standard deviations of the observable variables.

#### What does this tell you upfront about your ability to estimate price elasticities?

It would be easier at Central because the price varies more?

#### Repeat for Kraft

#### 4. Price plots: Construct time-series plots of sales and prices for Hellmans in the Central division and for Jewel (i.e. weeks on the X-axis, prices and unit-sales on the Y-axis). Repeat the exercise for Kraft. Describe the differences or similarities between Kraft and Hellman's pricing policies in each account.

```{r}
plot(hellman_at_jewel$date2, hellman_at_jewel$price, col = "blue", xy.labels = F, type = "l")

plot(hellman_at_central$date2, hellman_at_central$price, col = "red", type = 'l')

points(kraft_at_jewel$date2, kraft_at_jewel$price, col = "purple", type = 'l')

points(kraft_at_central$date2, kraft_at_central$price, col = "green", type = 'l')
```

They roughly follow similar rises and falls but it is hard.

#### Repeat the exercise for Kraft.

#### 5. Scatter-plots: Construct scatter-plots of sales versus prices for Hellmans in the Central division and for Jewel (i.e. prices on the Y-axis, unit-sales on the X-axis). Repeat the exercise for Kraft. Is there evidence for a negatively sloped demand-curve in the data? Eye-balling these plots, does demand appear more elastic in the Central Region or at Jewel (for either Hellman's or Kraft)?

```{r}
plot(kraft_at_jewel$sales_u, kraft_at_jewel$price, col = 'darkolivegreen3', xlim = c(0, max(kraft_at_central$sales_u)))
points(kraft_at_central$sales_u, kraft_at_central$price, col = 'darkorange2')
```

For Kraft it appears there is less demand elasticity at Central, where there appears to be stronger a negative correlation between price and total sales. However this effect is noticeable in both locations.

```{r}
plot(hellman_at_central$sales_u, hellman_at_central$price, col = 'darkolivegreen3')
points(kraft_at_central$sales_u, kraft_at_central$price, col = 'darkorange2')
```

I would say the reverse for Hellman. There is a noticeable negative trend between price and sales at Jewel for Hellman.

# Demand estimation

#### 1. Fit the "multiplicative" demand model discussed in class for Kraft and Hellman's at Jewel (i.e. 2 separate regressions, one for Hellman's, and one for Kraft).

```{r}
# Sales of Hellman at Jewel with Kraft price
#x = kraft_at_jewel, y = hellman_at_jewel
summary(lm(ln_sales_u ~ ln_price.x + ln_price.y, data = hellman.jewel_kraft.price))
```

Effect of kraft's price is positive with no significance. Effect of Hellman's price is negative with statistical significance. So an increase in price of Hellman's decreases sales of Hellman's.

```{r}
# Sales of Kraft at Jewel with Hellman price
#x = hellman_at_jewel, y = kraft_at_jewel
summary(lm(ln_sales_u ~ ln_price.x + ln_price.y, data = kraft.jewel_hellman.price))
```

Effect of hellman's price is positive with high significance (e.g. if hellman's is more expensive more people by kraft), effect of kraft is negative with strong significance.

#### 2. Elasticity differences: Is the demand elasticity higher (in absolute magnitude) at the Jewel account or in the Central Region? Can you offer some compelling explanations for the difference? (think of as many potential reasons as possible)

```{r}
hellman.central_kraft <- merged.df(kraft_at_central ,hellman_at_central)

# Sales of Hellman at Central  with Kraft price
#x = kraft_at_central, y = hellman_at_central
summary(lm(ln_sales_u ~ ln_price.x + ln_price.y, data = hellman.central_kraft))
```

Effect of both is negative but for kraft price the result is not significant.

```{r}
kraft.central_hellman.price <- merged.df(hellman_at_central, kraft_at_central)

# Sales of Kraft at Central with Hellman price
#x = hellman_at_central, y = kraft_at_central
summary(lm(ln_sales_u ~ ln_price.x + ln_price.y, data = kraft.central_hellman.price))
```

Effect of hellman's price on sales of Kraft is positive (e.g. if Hellman's is more expensive than sales of Kraft rise). Effect of hellman's price on sales of hellman is negative.


#### 3. Elasticity differences: Is the demand elasticity higher (in absolute magnitude) at the Jewel account or in the Central Region? Can you offer some compelling explanations for the difference? (think of as many potential reasons as possible)

Demand elasticity is greater for Hellman's overall. In other words, if Hellman's is more expensive, people will by Kraft but if Kraft is more expensive people are less likely to convert.

It appears demand for Hellman's is more elastic at the account level (Jewel) because increases in drive a decrease in sales by -4.5 not sure how to interpret.

Prices are more consistent across the account than accross the region?

#### 4. Forecasting demand under a price change: Using your regression results from the multiplicative demand model, compute the % change in unit sales for a 10% increase in the price of Kraft and Hellman's at Jewel. (Note: You can do this brute force in Excel, but for your benefit you should try to compute this on a sheet of paper with the help of a calculator).

```{r}
# Sales of Kraft at Jewel with Hellman price
#x = hellman_at_jewel, y = kraft_at_jewel
model <- lm(ln_sales_u ~ ln_price.x + ln_price.y, data = kraft.jewel_hellman.price)
# Coefficients:
# (Intercept)   ln_price.x   ln_price.y  
#       9.221        1.871       -4.439 
#To get the proportional change in Y associated with a p percent increase in X, calculate
# a = log([100 + p]/100) and take e^aβ
y <- model$coefficients
kraft_coeff <- y['ln_price.y'][[1]]
a <- log((100 + 10)/100)
kraft_effect <- exp(a*kraft_coeff)
prop_decrease <- log(kraft_effect)
# > kraft_effect
# decline in effect by 34%
# [1] 0.6550377

hellmans_coeff <- y['ln_price.x'][[1]]
a <- log((100 + 10)/100)
hellmans_effect <- exp(a*hellmans_coeff)
(prop_decrease <- log(hellmans_effect))
```

